---
title: "Structural Variation in *Drosophila santomea*"
author: "Brandon Turner"
date: "4/22/2021"
output: html_document
---
### Introduction

The appearance of new genes through the reorganization of existing genetic material is a unique source of genetic variation within populations. Chromosomal rearrangements act as a source of this genetic variation and divergence by shuffling DNA throughout the genome. Novel genetic variation in the forms of de novo gene formation, chimeric genes, or changes in gene expression can be introduced in the populations due to these mutations in genetic structure. To better understand the role that these various mutations play in evolutionary biology I observe changes in the *Drosophila* genome by analyzing 44 strains of *Drosophila santomea* and 22 strains of *Drosophila yakuba*. I identify 16,779 rearrangements through both populations and calculated the allele frequency of each population to assess population differentiation between *D. yakuba* and *D.santomea*. 

I speculate that **Hypothesis 1:** The distribution of the allele frequency for the rearrangement's is significantly different between the 3 populations, **Hypothesis 2:** In all populations, there is a positive correlation between the FST/deltap, where FST is calculated on Mainland Dyak and Island Dsant, and **Hypothesis 3 :** Deltap is significanlty different when comparing *D.sant* and *D.yak*.


### Methods

To calculate the allele frequencies for each subpopulation, I mapped the mutation call positions into the haplotype calls. After I had mapped the mutatation calls to their haplotype regions, I clustered the mutations using a clustering algorithim I wrote in Python. It clustered the mutations across strains if they were within 325 base pairs of each other because that is our sequence data's library insert size. Once the mutations were clustered across strains, I was able to calcualte the allele frequency for each rearrangment. Furthermore, by using regex I was able to calculate the allele frequency at a rearrangment site by subpopulation as well. The three subpopulations are island *D.santomea*, island *D.yakuba*, and mainland *D.yakuba*. 

Afterwards, I used BLAST to compare the sequence data at the rearrangment breakpoints (+/-1kb) against Repbase to search for matches to repetive elements, as well as the *Drosophila teisseire* reference genome so that I could polarize the anscetral state. If a rearrangmenet had one breakpoint match to a TE in Repbase it was marked as a TE insertion, and if both breakpoints matched it is assumed that the rearrangmenet is facilitating ectopic recombination. When polarizing the anscestral state, both breakpoints needed to map to the same location in the *D. tei* reference genome, and if they did then the allele frequency was flipped (1 - p_i; s.t p_i = allele frequency of i_th population).

I also used coverage data to correct for heterozygosity. By checking the coverage data and searching for 2x coverage at the breakpoints we can infer that the site is heterzygous, and when calculating allele frequency we consider the site 2/2 rather than 1/2. 

Once these corrections are added to the calculation of the allele frequency, I expored the data to R so that it could be plotted. Once I had the data in R, I ran statistical analysis to explore my hypotheses using the Wilcoxin rank-sum test, the Kolmogorov–Smirnov test, and the Kendall test. I also calculate and analyze deltap. Deltap = p_1 - p_2; s.t p_i = the allele frequency of the i_th populaton.

### Results

Figure 1 is a histogram of the allele frequencies for each subpopulation that has been log10 transformed.

```{r echo=FALSE, message=FALSE}
dsan_grey_fullFile = read.table('/Users/bturne48/Documents/GitHub/RogersLab/RemoteEdit/TempAsc/dsan_grey_deltap_plot.txt')
oran_grey_fullFile = read.table('/Users/bturne48/Documents/GitHub/RogersLab/RemoteEdit/TempAsc/oran_grey_deltap_plot.txt')
par(mfrow=c(1,3))
hist(log10(dsan_grey_fullFile$V9), main='Allele frequency: D.sant', xlab='Log10 allele freq')
hist(log10(dsan_grey_fullFile$V12), main='Allele frequency: Mainland D.yak', xlab='Log10 allele freq', sub='Figure 1')
hist(log10(oran_grey_fullFile$V9), main='Allele frequency: Island D.yak', xlab='Log10 allele freq')
```


**Hypothesis 1:** The results for all tests show that the distribtuion of allele frequencies between the populations for all tests are significantly different after running a Wilcoxin-rank sum test and a Kolmogorov-Smirnov test.

######*D.santomea* vs. Mainland *D.yakuba*:

```{r message=TRUE, warning=FALSE}
wilcox.test(dsan_grey_fullFile$V9, dsan_grey_fullFile$V12)
ks.test(dsan_grey_fullFile$V9, dsan_grey_fullFile$V12)
```
######Island *D.yakuba* vs. Mainland *D.yakuba*:

```{r message=TRUE, warning=FALSE}
wilcox.test(dsan_grey_fullFile$V9, dsan_grey_fullFile$V12)
ks.test(dsan_grey_fullFile$V9, dsan_grey_fullFile$V12)
```

**Hypothesis 2:** The Kendall test shows that the two populations are not independent of each other.

```{r}
library(Kendall)
Kendall(dsan_grey_fullFile$V9-dsan_grey_fullFile$V12, dsan_grey_fullFile$V15)
```

**Hypothesis 3 :** Deltap shows the differentiation of populations at sites accross the genome when comparing the (p_santome - p_main_yakuba) and between (p_island_yakuba - p_main_yakuba). The points highlighted in light green are rearrangements that are TE insertions, and dark green points are sites that are likely facilitating ectopic recombination. Figure 2-3/5-6 shows this for an autosome, and figure 4-5/7-8 shows this for the sex chromosome.

```{r echo=FALSE}
chrom = c('2L', 'X')
z = 2
for (i in 1:2) {

  # seperate background delta p (oran)
  oran_chrom1_background = oran_grey_fullFile[oran_grey_fullFile$V1==chrom[i] & is.na(oran_grey_fullFile$V9)!=TRUE & is.na(oran_grey_fullFile$V12)!=TRUE & oran_grey_fullFile$V16==0,]
  oran_chrom2_background = oran_grey_fullFile[oran_grey_fullFile$V4==chrom[i] & is.na(oran_grey_fullFile$V9)!=TRUE & is.na(oran_grey_fullFile$V12)!=TRUE & oran_grey_fullFile$V16==0,]
  # seperate background delta p (dsan)
  dsan_chrom1_background = dsan_grey_fullFile[dsan_grey_fullFile$V1==chrom[i] & is.na(dsan_grey_fullFile$V9)!=TRUE & is.na(dsan_grey_fullFile$V12)!=TRUE & dsan_grey_fullFile$V16==0,]
  dsan_chrom2_background = dsan_grey_fullFile[dsan_grey_fullFile$V4==chrom[i] & is.na(dsan_grey_fullFile$V9)!=TRUE & is.na(dsan_grey_fullFile$V12)!=TRUE & dsan_grey_fullFile$V16==0,]
  
  # seperate 1 sided delta p
  oran_chrom1_oneS = oran_grey_fullFile[oran_grey_fullFile$V1==chrom[i] & oran_grey_fullFile$V16==1,]
  dsan_chrom1_oneS = dsan_grey_fullFile[dsan_grey_fullFile$V1==chrom[i] & dsan_grey_fullFile$V16==1,]

  # seperate 2 sided delta p
  oran_chrom1_twoS = oran_grey_fullFile[oran_grey_fullFile$V1==chrom[i] & oran_grey_fullFile$V16==2,]
  oran_chrom2_twoS = oran_grey_fullFile[oran_grey_fullFile$V4==chrom[i] & oran_grey_fullFile$V16==2,]
  dsan_chrom1_twoS = dsan_grey_fullFile[dsan_grey_fullFile$V1==chrom[i] & dsan_grey_fullFile$V16==2,]
  dsan_chrom2_twoS = dsan_grey_fullFile[dsan_grey_fullFile$V4==chrom[i] & dsan_grey_fullFile$V16==2,]

  
  
  ### plots 
  
  ## dsan background w/one side
  par(mfrow=c(2,2))
  plot(((dsan_chrom1_background$V2+dsan_chrom1_background$V3)/2)/1000000, dsan_chrom1_background$V9-dsan_chrom1_background$V12,
       main=expression(paste(Delta, 'p for D.san TE insertion')), xlab=paste('Pos (MB) Chrom:',chrom[i]), ylab=expression(paste(Delta, 'p')),         pch=20,col=rgb(red=0.33, green=0.33, blue=0.33, alpha=0.5), ylim=c(-1,1), bty='n', sub=paste('Figure:', z))
  points(((dsan_chrom2_background$V5+dsan_chrom2_background$V6)/2)/1000000, dsan_chrom2_background$V9-dsan_chrom2_background$V12,
         main=paste('SVs: p_Dsan - p_Grey: ',chrom[i]), xlab='Pos (MB)', ylab='Delta p', pch=20,
         col=rgb(red=0.33, green=0.33, blue=0.33, alpha=0.5), ylim=c(-1,1))
  # one sided
  points(((dsan_chrom1_oneS$V2+dsan_chrom1_oneS$V3)/2)/1000000, dsan_chrom1_oneS$V9-dsan_chrom1_oneS$V12, pch=20, 
         col=rgb(red=0.1, green=0.66, blue=0.1, alpha=0.5))
  
  z = z+1
  
  ## dsan background w/two side
  plot(((dsan_chrom1_background$V2+dsan_chrom1_background$V3)/2)/1000000, dsan_chrom1_background$V9-dsan_chrom1_background$V12,
      main=expression(paste(Delta, 'p for D.san: ectopic recombination')), xlab=paste('Pos (MB) Chrom:',chrom[i]), ylab=expression(paste(Delta, 'p')), pch=20,
       col=rgb(red=0.33, green=0.33, blue=0.33, alpha=0.5), ylim=c(-1,1), bty='n', sub=paste('Figure:', z))
  points(((dsan_chrom2_background$V5+dsan_chrom2_background$V6)/2)/1000000, dsan_chrom2_background$V9-dsan_chrom2_background$V12,
         main=paste('SVs: p_Dsan - p_Grey: ',chrom[i]), xlab='Pos (MB)', ylab='Delta p', pch=20,
         col=rgb(red=0.33, green=0.33, blue=0.33, alpha=0.5), ylim=c(-1,1))
  # two sided
  points(((dsan_chrom1_twoS$V2+dsan_chrom1_twoS$V3)/2)/1000000, dsan_chrom1_twoS$V9-dsan_chrom1_twoS $V12, pch=20, 
         col=rgb(red=0.0, green=0.33, blue=0.0, alpha=0.5))
  points(((dsan_chrom2_twoS$V5+dsan_chrom2_twoS$V6)/2)/1000000, dsan_chrom2_twoS$V9-dsan_chrom2_twoS $V12, pch=20, 
         col=rgb(red=0.0, green=0.33, blue=0.0, alpha=0.5))

  z = z+1
  
  ## oran
  # one sided
  #par(mfrow=c(1,2))
  plot(((oran_chrom1_background$V2+oran_chrom1_background$V3)/2)/1000000, oran_chrom1_background$V9-oran_chrom1_background$V12,
       main=expression(paste(Delta, 'p for D.yak TE insertion')), xlab=paste('Pos (MB) Chrom:',chrom[i]), ylab=expression(paste(Delta, 'p')), pch=20,
       col=rgb(red=0.33, green=0.33, blue=0.33, alpha=0.5), ylim=c(-1,1), bty='n', sub=paste('Figure:', z))
  points(((oran_chrom2_background$V5+oran_chrom2_background$V6)/2)/1000000, oran_chrom2_background$V9-oran_chrom2_background$V12,
         main=paste('SVs: p_oran - p_Grey: ',chrom[i]), xlab='Pos (MB)', ylab='Delta p', pch=20,
         col=rgb(red=0.33, green=0.33, blue=0.33, alpha=0.5), ylim=c(-1,1))
  points(((oran_chrom1_oneS$V2+oran_chrom1_oneS$V3)/2)/1000000, oran_chrom1_oneS$V9-oran_chrom1_oneS$V12, pch=20, 
         col=rgb(red=0.1, green=0.66, blue=0.1, alpha=0.5))
  
  z = z+1

  # two sided
  plot(((oran_chrom1_background$V2+oran_chrom1_background$V3)/2)/1000000, oran_chrom1_background$V9-oran_chrom1_background$V12,
      main=expression(paste(Delta, 'p for D.san: ectopic recombination')), xlab=paste('Pos (MB) Chrom:',chrom[i]), ylab=expression(paste(Delta, 'p')), pch=20,
       col=rgb(red=0.33, green=0.33, blue=0.33, alpha=0.5), ylim=c(-1,1), bty='n', sub=paste('Figure:', z))
  points(((oran_chrom2_background$V5+oran_chrom2_background$V6)/2)/1000000, oran_chrom2_background$V9-oran_chrom2_background$V12,
         main=paste('SVs: p_oran - p_Grey: ',chrom[i]), xlab='Pos (MB)', ylab='Delta p', pch=20,
         col=rgb(red=0.33, green=0.33, blue=0.33, alpha=0.5), ylim=c(-1,1))
  points(((oran_chrom1_twoS$V2+oran_chrom1_twoS$V3)/2)/1000000, oran_chrom1_twoS$V9-oran_chrom1_twoS$V12, pch=20,
         col=rgb(red=0.0, green=0.33, blue=0.0, alpha=0.5))
  points(((oran_chrom2_twoS$V5+oran_chrom2_twoS$V6)/2)/1000000, oran_chrom2_twoS$V9-oran_chrom2_twoS$V12, pch=20,
         col=rgb(red=0.0, green=0.33, blue=0.0, alpha=0.5))
  
  print(wilcox.test(dsan_grey_fullFile$V9-dsan_grey_fullFile$V12, oran_grey_fullFile$V9-dsan_grey_fullFile$V12))
  
  z = z+1

}
```



### Discussion

**Hypothesis 1: ** The p-values of the Wilcoxin rank-sum test and Kolmogorov-Smirnov test tell us that the distributions of the allele frequencies between the subpopulations are significantly different from each other. This is helpful information because it tells us that chromosomal rearrangmenets are at higher or lower frequency in some populations, and could suggest that there is an over/underepresentation for some reason.

**Hypothesis 2: ** The p-value of the Kendall test tells us that we can reject the null hypothesis that the deltap calculated and the FST statistic are independent. With a tau correation coeffiecient of 0.327, we can infer that generally as the values of delta p increase, so does FST. Deltap and FST are both statistics that measure differentialtion between populations. The fact that the deltap plots from figures 2/3 show a postive skew, combined with the fact that there is a positive correlation between deltap and FST in this case tell us that these statistics that tell us similar information are agreeing each other. This is a good way to validate that my methods don't have data parsing errors. 

**Hypothesis 3: ** The p-values of the Wilcoxin rank-sum test show that the distribution of deltap is signifcantly different between the two populations. Deltap = p_1 - p_2, s.t postive values indicate that at a site where there was a higher allele frequency in p_1. 

Figure 3-8 shows that the differentation is more obvious to the eye on the sex chromosome, which is not unexpected as the mutation rate of the sex chromosome is generally faster in *Drosophila*. The plots show that generally the points that are green are more likely to be positive, meaning they had a higher allele frequency in *D.sant*. When comparing the deltap plots for the TE insertions, we can see that the positive skew is more visible when p_1 is *D.sant* rather than when p_1 is the island *D.yak*. This would seem to imply that *D.sant* have rearrangement sites that are more present in *D.sant* and are associated with TE's. Furthermore, perhaps based on knowledge of the populations, one could theorize that *D.sant* had a burst of TEs after moving to the island which helped the population adapt. This would explain why the deltap is higher when p_1 = *D.sant*, because *D.sant* have been on the island longer than the island *D.yak*.